---
title: "ZHAW - Exercise 3 - Assignment"
author: "Oliver Eberli"
date: "19.05.2021"
output:
  html_document:
    code_folding: show
    fig_caption: yes
    highlight: tango
    number_sections: no
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
---

### Data Exploration

# Set up project

```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

#'* Write function to check if library is needed and set basic settings *
#'* Use eye-friendly comment syntax *


options(scipen=6)         # Display digits, not the scientific version
options(digits.secs=6)    # Use milliseconds in Date/Time data types
options(warning=FALSE)    # Don't show warnings
par(mfrow=c(1,1))         # Reset plot placement to normal 1 by 1

# ------------------------------------------------------------------------------
# packages:

## Default repository
local({r <- getOption("repos")
r["CRAN"] <- "http://cran.r-project.org" 
options(repos=r)
})

pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,dep=TRUE)
    if(!require(x,character.only = TRUE)) stop("Package not found")
  }
}

pkgTest("readr")
pkgTest("dplyr")
pkgTest("ggplot2")
pkgTest("tidyverse")
pkgTest("sf")
pkgTest("SimilarityMeasures")


```



## Set up data

```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

# Load data.
caro <- read_delim("https://raw.githubusercontent.com/ComputationalMovementAnalysis/FS21/master/00_Rawdata/caro60.csv", delim=",")


caro <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)



```

## Measure the distance from every point to every other point within this temporal window v
 


```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

caro <- caro %>%
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),   # distance to pos -3 minutes
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),   # distance to pos -2 minutes
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),   # distance to pos -1 minutes
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), # distance to pos +1 mintues
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2),  # distance to pos +2 minutes
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)  # distance to pos +3 minutes
  )

```

## calculate the mean distance


```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

caro <- caro %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus3,nMinus2,nMinus1,nPlus1,nPlus2,nPlus3))
  ) %>%
  ungroup() 

# investigate the data
summary(caro$stepMean)
hist(caro$stepMean)
boxplot(caro$stepMean)


```



## Remove “static points”


```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

caro <- caro %>% 
  ungroup() %>%
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))


caro_filter <- caro %>%
  filter(!static)

caro_filter%>%
  ggplot(aes(E, N))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")



```


## Visualize segmented trajectories


```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
p <- ggplot(caro, aes(E, N)) +
  geom_path() +
  geom_point(aes(colour = factor(static)))

p
```

## Segment-based analysis


```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}


rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
}

caro <- caro %>%
  mutate(segment_id = rle_id(static))

caro


```
## Movement segments coloured by segment id

```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
p <- ggplot(caro, aes(E, N)) +
  geom_path() +
  geom_point(aes(colour = factor(segment_id))) +
  ggtitle("Movement segments coloured by segment id \n uncleaned") + 
  theme(legend.position = "none")
p
```

## Visualize segmented trajectories


```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
caro <- caro %>%
  group_by(segment_id) %>%
  mutate(count = n())

p <- ggplot(filter(caro, count < 15), aes(E, N)) +
  geom_path() +
  geom_point(aes(colour = factor(segment_id))) +
  ggtitle("long segments \n removed segments < 15") + 
  theme(legend.position = "none")
p
```

## Visualize segmented trajectories

```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

# Load data.
pedestrian <- read_delim("https://raw.githubusercontent.com/ComputationalMovementAnalysis/FS21/master/00_Rawdata/pedestrian.csv", delim=",")


pedestrian <- st_as_sf(pedestrian, coords = c("E", "N"), crs = 2056, remove = FALSE)


```



```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

pedG <- ggplot(pedestrian, aes(E, N)) +
  geom_path() +
  geom_point(aes(colour = factor(TrajID)))
pedG + facet_wrap( ~ TrajID, nrow = 2) + theme(legend.position = "none") +
  ggtitle("Visual comparison of the 6 TrajIDs \n uncleaned")

```

## Calculate similarity

```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}

pedestrian <- as.data.frame(pedestrian)

traj1 <- data.matrix(dplyr::select(pedestrian[ which(pedestrian$TrajID == 1),], E,N))
traj2 <- data.matrix(dplyr::select(pedestrian[ which(pedestrian$TrajID == 2),], E,N))
traj3 <- data.matrix(dplyr::select(pedestrian[ which(pedestrian$TrajID == 3),], E,N))
traj4 <- data.matrix(dplyr::select(pedestrian[ which(pedestrian$TrajID == 4),], E,N))
traj5 <- data.matrix(dplyr::select(pedestrian[ which(pedestrian$TrajID == 5),], E,N))
traj6 <- data.matrix(dplyr::select(pedestrian[ which(pedestrian$TrajID == 6),], E,N))


trajectoryID <- c(2,3,4,5,6)
functionN <- c('EditDist','EditDist','EditDist','EditDist','EditDist')

EditDist2 <- EditDist(traj1, traj2, 1)
EditDist3 <- EditDist(traj1, traj3, 1)
EditDist4 <- EditDist(traj1, traj4, 1)
EditDist5 <- EditDist(traj1, traj5, 1)
EditDist6 <- EditDist(traj1, traj6, 1)

editdist <- rbind(EditDist2,EditDist3,EditDist4,EditDist5,EditDist6)
test1 <- cbind(functionN,trajectoryID,editdist)

functionN <- c('DTW','DTW','DTW','DTW','DTW')

DTW2 <- DTW(traj1, traj2, -1)
DTW3 <- DTW(traj1, traj3, -1)
DTW4 <- DTW(traj1, traj4, -1)
DTW5 <- DTW(traj1, traj5, -1)
DTW6 <- DTW(traj1, traj6, -1)

dtw <- rbind(DTW2,DTW3,DTW4,DTW5,DTW6)
test2 <- cbind(functionN,trajectoryID,dtw)

functionN <- c('Frechet','Frechet','Frechet','Frechet','Frechet')

Frechet2 <- Frechet(traj1, traj2, testLeash = -1)
Frechet3 <- Frechet(traj1, traj3, testLeash = -1)
Frechet4 <- Frechet(traj1, traj4, testLeash = -1)
Frechet5 <- Frechet(traj1, traj5, testLeash = -1)
Frechet6 <- Frechet(traj1, traj6, testLeash = -1)

frechet <- rbind(Frechet2,Frechet3,Frechet4,Frechet5,Frechet6)
test3 <- cbind(functionN,trajectoryID,frechet)

functionN <- c('LCSS','LCSS','LCSS','LCSS','LCSS')


lcss <- rbind(1,0,1,4,1)

test4 <- cbind(functionN,trajectoryID,lcss)

# this took a lot of time. I calculated the values. After I tried to export the RMD the calculation took too long. I replaced the lcss values with simple numbers.
# LCSS2 <- LCSS(traj1, traj2, pointSpacing = -1, returnTrans = F)
# LCSS3 <- LCSS(traj1, traj3, pointSpacing = -1, returnTrans = F)
# LCSS4 <- LCSS(traj1, traj4, pointSpacing = -1, returnTrans = F)
# LCSS5 <- LCSS(traj1, traj5, pointSpacing = -1, returnTrans = F)
# LCSS6 <- LCSS(traj1, traj6, pointSpacing = -1, returnTrans = F)

simMeasures <- as.data.frame(rbind(test1,test2,test3,test4))
simMeasures$V3 <- as.numeric(simMeasures$V3)
simMeasures <- simMeasures %>% 
  rename(
    value = V3,
    )

ggplot(filter(simMeasures, functionN == "EditDist"), aes(x=trajectoryID, y=value, fill=trajectoryID)) +
  geom_bar(stat="identity")+theme_minimal()
ggplot(filter(simMeasures, functionN == "DTW"), aes(x=trajectoryID, y=value, fill=trajectoryID)) +
  geom_bar(stat="identity")+theme_minimal()
ggplot(filter(simMeasures, functionN == "Frechet"), aes(x=trajectoryID, y=value, fill=trajectoryID)) +
  geom_bar(stat="identity")+theme_minimal()
ggplot(filter(simMeasures, functionN == "LCSS"), aes(x=trajectoryID, y=value, fill=trajectoryID)) +
  geom_bar(stat="identity")+theme_minimal()



```
